---
title: 'PSYC259: Statistics in R Assignment'
author: "Amanda Mae Woodward"
date: "11/2/2020"
output: html_document
---
##Part 1
1. Open the diamonds dataset from ggplot2.   
```{r}
library(ggplot2)
data("diamonds")
```

2. Run a correlation between the x and y variables. What do you notice? 
```{r}
cor.test(diamonds$x, diamonds$y)
#there is a strong significant positive correlation. 
#The variables are essentially measuring the same thing. 
#There is also a very small CI. 
```

3. What is the correlation between price and carat?
```{r}
cor.test(diamonds$carat, diamonds$price)
#there is a strong, positive, significant relationship between carat and price 
#of the diamonds. The CI is also very tight. 
```

4. Create a linear model predicting price from table. What is the coefficient for table? 
```{r}
Q4lm <- lm(price ~ table, data = diamonds)
summary(Q4lm)
#226.984      
```

5. Run a regression and an Anova predicting price from color. What do you notice? 
```{r}
Q5lmreg <- lm(price ~color, data = diamonds)
summary(Q5lmreg)
library(tidyr)
library(car)
anova(lm(Q5lmreg))
anova(Q5lmreg)
Anova(lm(Q5lmreg))
Anova(Q5lmreg)

#well first of all, I don't recall going over 
#the differences of these ANOVA functions
#Buuuut, in ALL these models the F statistic is the same
```

6. Open the warpbreaks dataset. What test would you use to determine the correlation between wool type and breaks?
```{r}
data("warpbreaks")

class(warpbreaks$wool)
class(warpbreaks$breaks)
levels(warpbreaks$wool)
```

7. What is the correlation between wool type and breaks? 
```{r}
Q7cor <- glm(wool~breaks, data = warpbreaks, family = "binomial")
summary(Q7cor)
```
-0.03648: a nonsignificant weak, negative correlation


8. A student has a 25 question, multiple choice test. Each question has 4 choices. If they answered 17 questions correctly, did their responses differ from chance?  
```{r}

```






--------------------------
##Part 2
Here is a little information about the variables in the dataset you'll use for part 2. 
**School variables:**
schoolid: school's ID number
housepov: % of households in neighborhood of school below poverty level

**Classroom variables:**
classid: classroom's ID number
yearstea: years of teacher's experience in teaching the first grade
mathprep: the number of preparatory courses on the first grade contents and methods followed by teacher
mathknow: teacher's knowledge of first grade math contents

**Student variables:**
childid: Student ID
mathgain: student's gain in math achievement scores from spring of K through spring of first grade
mathkind: student's math score in spring of kindergarten year
sex: student's sex
minority: if minority or not
ses: child's socioeconomic status

1. Import the multilevel data from the SchoolsClassStudents.csv file
```{r}
Part2data <- read.csv("schoolsClassStudents.csv")
```

2. Scale and center your mathgain, ses, and mathprep variables.
```{r}
Part2data$Cmathgain <-scale(Part2data$mathgain)
Part2data$Cses <- scale(Part2data$ses)
Part2data$Cmathprep <- scale(Part2data$mathprep)
```

3. Load lme4 and lmerTest
```{r}
library(lme4)
library(lmerTest)
```

4. Run an intercept only model predicting math improvement for students.
```{r}
Q4Bmodel <- lmer(classid ~ 1 + (1 | Cmathgain), data = Part2data)
summary(Q4Bmodel)
```

BONUS: What is the between-class variance? What is the within-class variance?
between-class: 157.512      
within-class : 2.633  
?? IDK honestly

5. Run a model of students nested within classrooms to determine if the first-level predictor of child SES predicts child's improvement in scores, without modeling a random slope.
```{r}
Q5model<- lmer(Cmathgain ~ Cmathprep + (1 | classid), data = Part2data)
summary(Q5model)
```

BONUS: What does this tell us and what does a fixed slope not tell us?

6. Compare the model from level-1 predictor model of SES (question 5) to the intercept only model with a likelihood ratio test
```{r}

```

7. Run the level-1 predictor model of SES (question 5) but now model with a random slope for SES. Does your inference change? What parameter in the output is new with a random slope for SES?
```{r}

```

BONUS: You get a warning here. What is it and why might you be getting it?

8. Compare the level-1 predictor model of SES with a fixed slope and random intercept (question 5) to the intercept-only model (question 4). Now, compare the level-1 predictor model of SES with a random slope and random intercept (question 7) to the intercept-only model. How does your inference differ? What do these different LRTs tell you?
```{r}

```

9. Run a model to determine of students nested within classrooms to determine if the level 2 predictor of teacher's preparatory courses predict child's improvement in scores?
```{r}

```

10. Run a model with an interaction between student level SES and classroom level preparatory courses (i.e. cross-level interaction) predicting child math improvements, without modeling either as a random slope.
```{r}

```

11. Run a model with a cross-level interaction between student level SES and classroom level preparatory courses predicting child math improvement, while modeling SES as a random slope.
```{r}

```

BONUS: What looks different between the model in question eight versus the model in question nine? Why might this be?